---
title: "Week 7 Sampling Project Analysis"
format: pdf
editor: visual
---

## Optimum sub-sample size

```{r}
## c_n and c_m
c_n <- 3000
c_m <- 50

# roh for each variable 
roh_smoked <- 0.0306
roh_marijuana <- 0.0204
roh_age <- 0.0143
```

$$
m_{opt} = \sqrt{ C_n/C_m * 1-roh/roh }
$$

```{r}
### computing subsample size for each roh
mopt_smoked <- ceiling(sqrt((c_n/c_m) * ((1-roh_smoked)/roh_smoked)))
mopt_marijuana <- ceiling(sqrt((c_n/c_m) * ((1-roh_marijuana)/roh_marijuana)))
mopt_age <- ceiling(sqrt((c_n/c_m) * ((1-roh_age)/roh_age)))

# Print the optimum subsample sizes
cat("Optimum subsample size for Smoked: ", mopt_smoked, "\n")
cat("Optimum subsample size for Marijuana: ", mopt_marijuana, "\n")
cat("Optimum subsample size for Age: ", mopt_age, "\n")

```

$$
Samplingcost = n * c_n + n * m * c_m
$$

```{r}
### corresponding optimal number of first stage clusters, given the total budget 
sampling_cost <- 500000
firstcluster_smoked <- sampling_cost/(c_n + (c_m * mopt_smoked))
firstcluster_marijuana <- sampling_cost/(c_n + (c_m * mopt_marijuana))
firstcluster_age <- sampling_cost/(c_n + (c_m * mopt_age))

# Print the optimal number of clusters for each variable
cat("Optimal number of clusters for Smoked: ", firstcluster_smoked, "\n")
cat("Optimal number of clusters for Marijuana: ", firstcluster_marijuana, "\n")
cat("Optimal number of clusters for Age: ", firstcluster_age, "\n")
```

## Selecting overall sample size

```{r}
## Choosing to use the mean of the sample sizes
m_opt_overall <- 44
n_opt <- sampling_cost / (c_n + m_opt_overall * c_m)
n_total <- n_opt * m_opt_overall

cat("Optimal number of subsample size (m):", m_opt_overall, "\n")
cat("Optimal number of clusters (n_opt):", n_opt, "\n")
cat("Overall sample size (n_total):", n_total, "\n")

```

## Design effect and sampling variance

$$
deff_{new} = 1+(m_{new} -1) * roh
$$

```{r}
deff_smoked <- 1 + (m_opt_overall - 1) * roh_smoked
deff_marijuana <- 1 + (m_opt_overall - 1) * roh_marijuana
deff_age <- 1 + (m_opt_overall - 1) * roh_age

cat("Design effect for smoked:", deff_smoked, "\n")
cat("Design effect for marijuana:", deff_marijuana, "\n")
cat("Design effect for age:", deff_age, "\n")

```

$$
var(SRS) = p(1-p)/nm-1
$$

```{r}
# Element Variance 
var_ever_smoked <- 0.188
var_ever_marijuana <- 0.128
var_age_first <- 1

new_varsrs_smoked <- var_ever_smoked/(n_total-1)
new_varsrs_marijuana <- var_ever_marijuana/(n_total-1)
new_varsrs_age <- var_age_first/(n_total-1)  


# Print new SRS variances
cat("New SRS Variance for Smoked: ", new_varsrs_smoked, "\n")
cat("New SRS Variance for Marijuana: ", new_varsrs_marijuana, "\n")
cat("New SRS Variance for Age: ", new_varsrs_age, "\n")
```

$$
var_{cluster-new} (\bar{y}) = var_{srs} * deff_{new}
$$

```{r}
## New cluster design 
varcluster_smoked <- new_varsrs_smoked * deff_smoked 
varcluster_marijuana <- new_varsrs_marijuana * deff_marijuana 
varcluster_age <- new_varsrs_age * deff_age

cat("Expected Cluster Sample Variance for Smoked: ", varcluster_smoked, "\n")
cat("Expected Cluster Sample Variance for Marijuana: ", varcluster_marijuana, "\n")
cat("Expected Cluster Sample Variance for Age: ", varcluster_age, "\n")
```

```{r}
## Sampling Fraction and FPC
N <- 830138

sampling_fraction <- n_total/N
FPC <- 1 - sampling_fraction

# Print results
cat("Sampling fraction (f):", sampling_fraction, "\n")
cat("Finite Population Correction (FPC):", FPC, "\n")
```

## Justification for choosing subsample

```{r}
## Justification for choosing subsample

precision_smoked <- sqrt(varcluster_smoked)
precision_marijuana <- sqrt(varcluster_marijuana)
precision_age <- sqrt(varcluster_age)

cat("Precision for Smoked:", precision_smoked, "\n")
cat("Precision for Marijuana:", precision_marijuana, "\n")
cat("Precision for Age:", precision_age, "\n")
```

Inference:

In week7, we focused on refining our cluster sampling design to optimize both cost and statistical precision. By analyzing the intraclass correlation coefficients (ICCs) for three key variables—ever smoked (0.0306), ever used marijuana (0.0204), and age at first approach (0.0143), we calculated the optimal number of students to sample within each cluster. 

This resulted in subsample sizes of 44, 54, and 65 students, respectively. These figures balance the trade-off between the costs of sampling additional students and the benefits of increased precision.​

Considering budget constraints, we determined that sampling 44 students per cluster would allow us to include approximately 96 clusters, totaling around 4,231 students. This configuration ensures that we stay within our $500,000 budget while maintaining a manageable cluster size.​

We then calculated the design effects for each variable using the formula:​

Design Effect=1+(Cluster Size−1)×ICC


This yielded design effects of 2.32 for smoking, 1.88 for marijuana use, and 1.61 for age at first approach. These values indicate the extent to which the variance is inflated due to the clustering in our sampling design.​

Subsequently, we computed the variances under both simple random sampling (SRS) and our cluster sampling design. For instance, the SRS variance for smoking was approximately 0.000044, which increased to 0.000103 under the cluster design. This increase is expected due to the design effect.​

The sampling fraction, calculated as the total sample size divided by the population size (4,231 / 830,138), was approximately 0.0051. This small fraction results in a finite population correction (FPC) factor of about 0.9949, indicating a minimal impact on variance estimates.​

Finally, we assessed the precision of our estimates by taking the square root of the variances under the cluster design. The precisions were approximately 0.0101 for smoking, 0.0075 for marijuana use, and 0.0195 for age at first approach. These values suggest that our sampling design achieves a reasonable balance between cost and statistical accuracy.​


