---
title: "Week 11"
format: pdf
editor: visual
---

## Step 1: Load the data and examine the school roster

```{r}
library(readxl)
library(dplyr)
library(stringr)

raw_data <- read_excel("sample_school_student_list.xls", col_names = FALSE)

colnames(raw_data) <- c("Col1", "Col2", "Col3")

student_rows <- raw_data %>%
  filter(!is.na(Col1), !is.na(Col2), !is.na(Col3)) %>%  
  filter(str_detect(Col1, "^[78]$"),                    
         str_detect(Col2, "^[A-Z]"),                    
         !str_detect(Col1, "Grade|Teacher|Surname"))    

# Count total students
actual_school_size <- nrow(student_rows)
print(paste("Actual school size:", actual_school_size))

# Count students by grade
grade_counts <- student_rows %>%
  group_by(Col1) %>%
  summarize(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

# Rename columns for clarity
colnames(grade_counts) <- c("Grade", "Count", "Percentage")
print(grade_counts)

# Add student IDs for reference
student_rows <- student_rows %>%
  mutate(student_id = row_number())

# Preview the data
head(student_rows)
```

## Actual Size of the Sampled Middle School
The actual size of the sampled middle school, based on the classroom rosters, is $M_{actual} = 219$.

## Sampling Rate to Achieve EPSEM

So to achieve EPSEM in a two-stage PPeS sampling design, the second-stage sampling rate ($f_{2nd}$) applied within the sampled school needs to satisfy the following relationship

$242 \times f_{2nd} = m^*_7$

Therefore, the sampling rate that you would apply to this school to achieve EPSEM is:
$\boxed{f_{2nd} = \frac{m^*_7}{242}}$


```{r}
# Calculating the sampling rate for EPSEM
# Given parameters
mos_school <- 242  # Measure of size for this school from frame
actual_school_size <- 219  # Actual number of students from roster
m_star_h <- 43  # Sampling interval for Region 7 

# Calculate the sampling rate to maintain EPSEM
sampling_rate <- m_star_h / mos_school
print(paste("Within-school sampling rate:", round(sampling_rate, 4)))

print(paste("Sampling rate value:", sampling_rate))
print(paste("Sampling interval (k) =", 1/sampling_rate))
```

## Implementing systematic sampling for student selection

```{r}
# First, calculate the expected sample size from this school
expected_sample_size <- actual_school_size * (sampling_rate / mos_school)
print(paste("Expected sample size:", round(expected_sample_size, 2), "students"))

# Since sampling rate > 1 doesn't make sense for within-school sampling,
# we need to correctly interpret what this means:
# - The sampling rate is actually telling us we need to select ~38.59 students
# - This is approximately 1/6 of the total students (219)
# - So our within-school sampling rate should be around 38.59/219 ≈ 0.176

# Calculate the correct within-school sampling probability
within_school_prob <- min(1, n_opt / actual_school_size)
where_n_opt <- sampling_rate
print(paste("Adjusted within-school sampling probability:", round(within_school_prob, 4)))

# Function for systematic random sampling
systematic_sample <- function(data, expected_sample_size) {
  # Total number of students
  N <- nrow(data)
  
  # Calculate sampling interval
  k <- N / expected_sample_size
  
  # Select a random start between 1 and k
  set.seed(625)  # For reproducibility
  random_start <- runif(1, min = 1, max = k)
  
  # Generate selection numbers
  selection_numbers <- seq(from = random_start, to = N, by = k)
  
  # Round down to integer positions
  selection_indices <- floor(selection_numbers)
  
  # Ensure no duplicates and valid indices
  selection_indices <- unique(selection_indices)
  selection_indices <- selection_indices[selection_indices <= N]
  
  # Select the sample - sampling WITHOUT replacement
  sampled_students <- data[selection_indices, ]
  
  return(list(
    sample = sampled_students,
    random_start = random_start,
    interval = k,
    selection_indices = selection_indices
  ))
}

# Apply systematic sampling with the expected sample size
sampling_result <- systematic_sample(student_rows, 38.59)

# Print sample details
print(paste("Random start:", round(sampling_result$random_start, 2)))
print(paste("Sampling interval:", round(sampling_result$interval, 2)))
print(paste("Actual sample size:", nrow(sampling_result$sample), "students"))
```


## Analyzing the selected sample

```{r}
# Sample summary by grade
sample_summary <- sampling_result$sample %>%
  group_by(Col1) %>%
  summarize(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

# Rename columns for clarity
colnames(sample_summary) <- c("Grade", "Count", "Percentage")
print(sample_summary)

# Compare with population distribution
comparison <- data.frame(
  Grade = grade_counts$Grade,
  Population_Count = grade_counts$Count,
  Population_Pct = grade_counts$Percentage,
  Sample_Count = sample_summary$Count,
  Sample_Pct = sample_summary$Percentage
)
print(comparison)

# Create a detailed selection table
selection_details <- data.frame(
  Student_ID = sampling_result$sample$student_id,
  Grade = sampling_result$sample$Col1,
  Surname = sampling_result$sample$Col2,
  Given_Name = sampling_result$sample$Col3,
  Selection_Number = seq_along(sampling_result$selection_indices),
  Selection_Position = sampling_result$selection_indices
)

# Display the first few selected students
head(selection_details, 10)
```

## Now to verify EPSEM property

```{r}
# Calculate the selection probabilities
first_stage_prob <- mos_school / sum_mos_region7  # Would need total MOS for region 7
second_stage_prob <- nrow(sampling_result$sample) / actual_school_size
overall_intended_prob <- 1/m_star_h

# Expected sample size calculation
expected_sample_size <- actual_school_size * (sampling_rate / mos_school)

# Create verification table
verification <- data.frame(
  Parameter = c(
    "School MOS", 
    "Actual school size", 
    "Sampling interval (m*_7)", 
    "Intended within-school sampling rate",
    "Expected sample size",
    "Actual sample size",
    "Effective sampling rate",
    "Overall selection probability (1/m*_7)"
  ),
  Value = c(
    mos_school,
    actual_school_size,
    m_star_h,
    sampling_rate / mos_school,
    expected_sample_size,
    nrow(sampling_result$sample),
    nrow(sampling_result$sample) / actual_school_size,
    overall_intended_prob
  )
)

knitr::kable(verification, 
             digits = 6,
             caption = "Verification of EPSEM Property")
```

