---
title: "Week 10"
format: pdf
editor: visual
---

# Week 10- Tasks

1. Given your overall $m_{opt}, n_{opt}$ and N (based on the sampling frame), you’ve already
computed the overall sampling fraction, 𝑓. For each of the nine strata, compute the
required number of students to subsample from each sampled school based on the
stratified PPeS design **in order to maintain epsem across all strata.**

2. Do each of the schools that you sampled in a given region have the minimum sufficient
size, given the stratum-specific subsample sizes computed in Task #1? Do subsequent
schools on the list have the minimum sufficient size? If not, what will you do?

3. Begin to describe how you will physically select the subsample of students within a given
sampled school (or set of linked schools). What will your second-stage sampling rate be for a
given school within a given stratum? How will you acquire the updated rosters from each
school? What technique will you use to select the sample at the specified second-stage
rate?

4. Write down the overall sampling fraction based on the stratified PPeS design, indicating
the overall probability of inclusion for a given student, from a given school (or linked set of
schools), in a given stratum. Be careful with notation. Keep in mind that the MOS values
used for the sampled schools at the first stage and the denominator at the second stage (Did
you sample a single school? Or a linked set of schools?) will depend on your response to Task
#2 above.

```{r}
library(knitr)
n_opt <- 86.95652
m_opt <- 55
N <- 830138
sampling_fractions <- c(0.3043478, 0.5739130, 0.8434782, 0.5043478, 1.5391304, 3.2608695, 18.5826083, 18.6956518, 39.2521731)
f <- n_opt / N
required_subsample_sizes <- sampling_fractions * f * m_opt

# Compute allocated clusters
allocated_clusters <- ceiling((n_opt * proportions))

regions <- c(1:9)
total_students <- c(2893, 5474, 8044, 4855, 14692, 31140, 177421, 178447, 374755)
num_schools <- c(16, 30, 30, 31, 69, 123, 598, 517, 923)
proportions <- c(0.0035, 0.0066, 0.0097, 0.0058, 0.0177, 0.0375, 0.2137, 0.2150, 0.4514)

region_data <- data.frame(
  Region = regions,
  Total_Students = total_students,
  Num_Schools = num_schools,
  Proportion = proportions,
  Allocated_Clusters = round(allocated_clusters, 2),
  Sampling_Fraction = round(sampling_fractions, 4),
  Required_Subsample = required_subsample_sizes
)
kable(region_data)
```


```{r}
library(knitr)
n_opt <- 86.95652  # Optimal cluster size (students per cluster)
m_opt <- 55  # Optimal number of clusters
N <- 830138  # Total population size

# These appear to be the number of PSUs (schools) allocated to each stratum
sampling_fractions <- c(0.3043478, 0.5739130, 0.8434782, 0.5043478, 1.5391304, 3.2608695, 18.5826083, 18.6956518, 39.2521731)

# Overall sampling fraction
f <- (n_opt * m_opt) / N  # This should be (total sample size)/(population size)

# Calculate required subsample sizes - the issue is in this calculation
# The correct formula should use the sampling interval for each stratum (m*_h)
# For each stratum h, m*_h = Total_Students_h / allocated_clusters_h

regions <- c(1:9)
total_students <- c(2893, 5474, 8044, 4855, 14692, 31140, 177421, 178447, 374755)
num_schools <- c(16, 30, 30, 31, 69, 123, 598, 517, 923)
proportions <- c(0.0035, 0.0066, 0.0097, 0.0058, 0.0177, 0.0375, 0.2137, 0.2150, 0.4514)

# Compute allocated clusters correctly
allocated_clusters <- ceiling(m_opt * proportions)

# Calculate sampling intervals for each stratum
sampling_intervals <- total_students / allocated_clusters

# Calculate required subsample sizes
# For EPSEM across strata, required subsample = n_opt
# But we need to adjust by the ratio of MOS to actual size
# If MOS equals actual size, then required subsample = n_opt
required_subsample_sizes <- n_opt  # Same for all strata to maintain EPSEM

region_data <- data.frame(
  Region = regions,
  Total_Students = total_students,
  Num_Schools = num_schools,
  Proportion = proportions,
  Allocated_Clusters = allocated_clusters,
  Sampling_Interval = round(sampling_intervals, 1),
  Required_Subsample = round(required_subsample_sizes, 2)
)
kable(region_data)
```


The table you have provided shows the results of your calculations for the nine strata. Based on your code, the "Required_Subsample" for each stratum was computed using the formula:
Required Subsample$h$ = Sampling Fraction${opt}$ / N) × m${opt}$
Where:
•
Sampling Fraction$_h$ is the value provided in the "Sampling_Fraction" column for each stratum.
•
n$_{opt}$ = 86.95652
•
N = 830138
•
m$_{opt}$ = 55
Your code correctly implements this formula. However, the crucial point is whether this formula and the "Sampling_Fraction" values are appropriate for determining the required subsample size to maintain epsem across all strata in a stratified PPeS design.