---
title: "Week 10"
format: pdf
editor: visual
---

# Week 10- Tasks

1.  Given your overall $m_{opt}, n_{opt}$ and N (based on the sampling frame), you’ve already computed the overall sampling fraction, 𝑓. For each of the nine strata, compute the required number of students to subsample from each sampled school based on the stratified PPeS design **in order to maintain epsem across all strata.**

2.  Do each of the schools that you sampled in a given region have the minimum sufficient size, given the stratum-specific subsample sizes computed in Task #1? Do subsequent schools on the list have the minimum sufficient size? If not, what will you do?

3.  Begin to describe how you will physically select the subsample of students within a given sampled school (or set of linked schools). What will your second-stage sampling rate be for a given school within a given stratum? How will you acquire the updated rosters from each school? What technique will you use to select the sample at the specified second-stage rate?

4.  Write down the overall sampling fraction based on the stratified PPeS design, indicating the overall probability of inclusion for a given student, from a given school (or linked set of schools), in a given stratum. Be careful with notation. Keep in mind that the MOS values used for the sampled schools at the first stage and the denominator at the second stage (Did you sample a single school? Or a linked set of schools?) will depend on your response to Task #2 above.

```{r}
library(knitr)
n_opt <- 96
m_opt <- 44
N <- 830138
proportions<- c(0.0035, 0.0066, 0.0097, 0.0058, 0.0177, 0.0375, 0.2137, 0.2150, 0.4514)
sampling_fractions <- c(0.3043478, 0.5739130, 0.8434782, 0.5043478, 1.5391304, 3.2608695, 18.5826083, 18.6956518, 39.2521731)
allocated_clusters <- ceiling((n_opt * proportions))

final_sample <- read_csv("final_sample.csv")

total_students <- aggregate(tot_all ~ Region, data = final_sample, sum)

f <- 0.005

required_subsample_sizes <- ceiling(total_students*f/allocated_clusters)

region_data <- data.frame(
  Region = regions,
  Total_Students = total_students,
  Num_Schools = num_schools,
  Proportion = proportions,
  Allocated_Clusters = round(allocated_clusters, 2),
  Sampling_Fraction = round(sampling_fractions, 4),
  Required_Subsample = required_subsample_sizes
)
kable(region_data)
```

```{r}
m_star <- n_opt/sum(allocated_clusters)
print(m_star)
nh <- total_students * f
m_star_regions <- nh/ allocated_clusters
print(ceiling(m_star_regions))
```

# Task 2 & 3

```{r}
library(readxl)
library(dplyr)

# Step 1: Load the school data
school_data <- read_csv("final_sample.csv")

# Required subsample sizes from Task #1
required_subsample_sizes <- c(1, 1, 1, 1, 2, 3, 3, 2, 7)

# Step 2: Loop through each region for Task 2
for (region in unique(school_data$Region)) {
  # Filter schools in the current region
  schools_in_region <- school_data[school_data$Region == region, ]
  
  # Get the required subsample size for the region
  required_size <- required_subsample_sizes[region]
  
  # Identify failing schools
  failing_schools <- schools_in_region[schools_in_region$tot_all < required_size, ]
  if (nrow(failing_schools) > 0) {
    cat("Some schools in Region", region, "do not meet the required subsample size.\n")
    
    # Combine failing schools
    combined_size <- sum(failing_schools$tot_all)
    combined_school <- failing_schools[1, ]  # Use the first failing school as a template
    combined_school$BNAME <- "Combined Failing Schools"
    combined_school$tot_all <- combined_size
    
    # Remove failing schools and add the combined unit
    schools_in_region <- schools_in_region[schools_in_region$tot_all >= required_size, ]
    schools_in_region <- rbind(schools_in_region, combined_school)
  }
  
  # Calculate second-stage sampling rate for each school
  schools_in_region <- schools_in_region %>%
    mutate(Second_Stage_Rate = required_size / tot_all)
  
  # Print the second-stage sampling rate
  cat("Region:", region, "\n")
  print(schools_in_region[, c("BNAME", "tot_all", "Second_Stage_Rate")])
  
  # Save the updated schools_in_region for Task 3
  assign(paste0("region_", region, "_data"), schools_in_region)
}

# Task 3: Perform systematic sampling for each school
for (region in unique(school_data$Region)) {
  # Load the updated schools_in_region from Task 2
  schools_in_region <- get(paste0("region_", region, "_data"))
  
  # Get the required subsample size for the region
  required_size <- required_subsample_sizes[region]
  
  # Perform systematic sampling for each school
  for (i in 1:nrow(schools_in_region)) {
    school_name <- schools_in_region$BNAME[i]
    total_students <- schools_in_region$tot_all[i]
    m_hi <- required_size  # Required subsample size for the school
    
    # Check if the total number of students is sufficient
    if (total_students < m_hi) {
      cat("Warning: Total students in school", school_name, "is less than the required subsample size.\n")
      next
    }
    
    # Generate a mock roster for the school (replace with actual roster data if available)
    roster <- data.frame(
      Student_ID = 1:total_students
    )
    
    # Perform systematic sampling
    k <- floor(total_students / m_hi)  # Sampling interval
    set.seed(625)  # For reproducibility
    random_start <- sample(1:k, 1)  # Random start
    sampled_indices <- seq(random_start, total_students, by = k)
    sampled_students <- roster[sampled_indices, ]
    
    # Print sampled students
    cat("Sampled students for school:", school_name, "\n")
    print(sampled_students)
  }
}
```

# Task 4

```{r}
# Step 1: Define the data
regions <- c(1:9)

compute_MOS_and_region_count <- function(school_data) {
  school_data %>%
    group_by(Region) %>%
    summarise(
      MOS = sum(tot_all) 
    )
}
result <- compute_MOS_and_region_count(ordered_by_tot_all)
print(result)

total_students <- result$MOS  # N_h
allocated_clusters <- c(1, 1, 1, 1, 2, 4, 21, 21, 44)  # Number of schools sampled per region
required_subsample_sizes <- c(1, 1, 1, 1, 2, 3, 3, 2, 7)  # m_hi

# Step 2: Compute first-stage probability (π_h^(1))
mos <- total_students  # Measure of size (MOS) for each region
sum_mos <- sum(mos)  # Total MOS across all regions
pi_h1 <- mos / sum_mos  # First-stage probability

# Step 3: Compute second-stage probability (π_hi^(2))
pi_h2 <- required_subsample_sizes / mos  # Second-stage probability

# Step 4: Compute overall probability of inclusion (π_hi)
pi_hi <- pi_h1 * pi_h2  # Overall probability of inclusion

# Step 5: Compute overall sampling fraction (f_h)
f_h <- required_subsample_sizes / total_students  # Overall sampling fraction

# Step 6: Create a data frame to display results
results <- data.frame(
  Region = regions,
  Total_Students = total_students,
  First_Stage_Prob = round(pi_h1, 6),
  Second_Stage_Prob = round(pi_h2, 6),
  Overall_Prob = round(pi_hi, 6),
  Sampling_Fraction = round(f_h, 6)
)

# Print the results
print(results)
```

inference:

