---
title: "Week 6 Sampling Project Analysis"
format: pdf
editor: visual
---

## Initial Parameters
```{r}
# Design effects from previous study
deff_smoked <- 2.5    # design effect for ever smoked cigarette
deff_marijuana <- 2.0 # design effect for ever smoked marijuana
deff_age <- 1.7      # design effect for age when first asked

# Previous study parameters
n_students <- 7500    # total sample size
n_clusters <- 150     # number of clusters (schools)

# Average cluster size
b_bar <- n_students / n_clusters
cat("Average cluster size (b-bar):", b_bar, "\n")

# Cost parameters
total_budget <- 500000  # total budget
cost_per_cluster <- 3000 # cost per school
cost_per_student <- 50   # cost per completed questionnaire
```

## Calculate Synthetic roh Values

Using the formula: roh = (deff - 1)/(b_bar - 1)

```{r}
# Function to calculate roh
calculate_roh <- function(deff, b_bar) {
  roh <- (deff - 1)/(b_bar - 1)
  return(roh)
}

# Calculate roh for each variable
roh_smoked <- calculate_roh(deff_smoked, b_bar)
roh_marijuana <- calculate_roh(deff_marijuana, b_bar)
roh_age <- calculate_roh(deff_age, b_bar)

# Create results table
roh_results <- data.frame(
  Variable = c("Ever Smoked", "Ever Marijuana", "Age First Asked"),
  Design_Effect = c(deff_smoked, deff_marijuana, deff_age),
  Roh = c(roh_smoked, roh_marijuana, roh_age)
)

knitr::kable(roh_results, 
             digits = 4,
             caption = "Synthetic roh Values by Variable")
```

## Optimal Design Calculations
We'll use the calculated roh values to determine the optimal cluster size and number of clusters given our budget constraints.

```{r}
library(tidyverse)

# We'll use the largest roh (Ever Smoked) to be conservative
roh_max <- max(roh_smoked, roh_marijuana, roh_age)

# Function to calculate optimal cluster size given roh and cost parameters
# Formula: n_opt = sqrt((cost_per_cluster)/(cost_per_student * roh))
calculate_optimal_cluster_size <- function(cost_cluster, cost_student, roh) {
  n_opt <- sqrt(cost_cluster / (cost_student * roh))
  return(n_opt)
}

# Calculate optimal cluster size
n_opt <- calculate_optimal_cluster_size(cost_per_cluster, cost_per_student, roh_max)
n_opt_rounded <- round(n_opt)

# Calculate how many clusters we can afford
# total_budget = (cost_per_cluster + cost_per_student * n_opt) * m_opt
# Solve for m_opt
m_opt <- total_budget / (cost_per_cluster + cost_per_student * n_opt_rounded)
m_opt_rounded <- floor(m_opt) # We round down to stay within budget

# Calculate total sample size
total_sample <- n_opt_rounded * m_opt_rounded

optimal_design <- data.frame(
  Parameter = c("Optimal Cluster Size (n_opt)", 
                "Rounded Cluster Size",
                "Optimal Number of Clusters (m_opt)",
                "Rounded Number of Clusters",
                "Total Sample Size",
                "Total Cost",
                "Average roh",
                "Expected Design Effect"),
  Value = c(n_opt,
            n_opt_rounded,
            m_opt,
            m_opt_rounded,
            total_sample,
            (cost_per_cluster + cost_per_student * n_opt_rounded) * m_opt_rounded,
            mean(c(roh_smoked, roh_marijuana, roh_age)),
            1 + (n_opt_rounded - 1) * roh_max)
)

knitr::kable(optimal_design, 
             col.names = c("Parameter", "Value"),
             digits = c(0, 2),
             caption = "Optimal Design Parameters")
```

## Cost Analysis

Calculating maximum possible number of clusters and students given budget constraints:

```{r}
# Function to calculate total cost
calculate_total_cost <- function(n_clusters, students_per_cluster) {
  cluster_costs <- n_clusters * cost_per_cluster
  student_costs <- n_clusters * students_per_cluster * cost_per_student
  total_cost <- cluster_costs + student_costs
  return(total_cost)
}

# Function to calculate maximum clusters and students
calculate_max_samples <- function(budget, cost_cluster, cost_student) {
  # Maximum possible clusters if all budget spent on clusters
  max_clusters <- floor(budget/cost_cluster)
  
  # Maximum possible students if all budget spent on students
  max_students <- floor(budget/cost_student)
  
  return(list(max_clusters = max_clusters, 
              max_students = max_students))
}

max_samples <- calculate_max_samples(total_budget, cost_per_cluster, cost_per_student)

cat("Maximum possible clusters (if only clusters):", max_samples$max_clusters, "\n")
cat("Maximum possible students (if only students):", max_samples$max_students, "\n")
```

## Summary Table

```{r}
summary_df <- data.frame(
  Parameter = c("Average Cluster Size (b-bar)", 
                "Total Budget",
                "Cost per Cluster",
                "Cost per Student",
                "Maximum Possible Clusters",
                "Maximum Possible Students",
                "Optimal Cluster Size",
                "Optimal Number of Clusters",
                "Total Sample Size"),
  Value = c(b_bar,
            total_budget,
            cost_per_cluster,
            cost_per_student,
            max_samples$max_clusters,
            max_samples$max_students,
            n_opt_rounded,
            m_opt_rounded,
            total_sample)
)

knitr::kable(summary_df, 
             col.names = c("Parameter", "Value"),
             digits = 0,
             caption = "Summary of Key Parameters")
```


Inferences:

The results highlight how intra-cluster correlation (ICC) directly shapes sampling efficiency. With ICC values ranging from 0.014 (Age First Asked) to 0.031 (Ever Smoked), schools exhibit moderate homogeneity in student behaviors—particularly for smoking. 

This clustering effect amplifies the required sample size, as seen in the design effect of 2.32, meaning nearly twice as many students are needed compared to a simple random sample.

Balancing cost and precision, the optimal design selects 96 schools with 44 students each (totaling 4,224 students). This strikes a middle ground between two extremes i.e. surveying 166 schools (maximizing geographic spread but undersampling students) or 10,000 students (ignoring school-level costs entirely). Not only that but the choice reflects real-world constraints like school recruitment expenses ($2,500 per school) and per-student data collection costs ($100), while prioritizing statistical power.

Now the reason this works is:

* Higher ICCs (e.g., smoking behavior) demand more clusters to capture between-school variability.
* The 44-student cluster size minimizes variance inflation from within-school similarities.
* Budget allocation mirrors practices in education research, where fixed cluster costs often dominate.

Practical considerations:
* Non-response adjustments helps in building a 10-15% buffer to ensure final samples meet power requirements.
* Parameter justification shows that benchmarking against prior studies (e.g., similar school-based smoking surveys) strengthens methodological credibility.
* Cost transparency helps in explicitly separating cluster/student costs helps stakeholders understand tradeoffs.
